Improving Item-Item Similarity Estimation during Collaborative Filtering

### 2014-03-25 paper/paper.tex
Tweaking techniques of predictivity model

The parameter values $\mu_s$ and $\sigma_s$ are obtained by fitting the
distribution of $GoldSimRand$ values in the training set (see Fig.
\ref{fig:Histograms}). For example, the $N \geq 80$ dataset we have $\mu_s =
0.27$ and $\sigma_s = 0.25$. The choice of $\mu_r$ and $\sigma_r$ is less clear.
Some insight can be gained, however, by examining the prior distribution of
ratings for our datasets, which are shown in Fig. \ref{fig:RatingDist}.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{RatingDist_40.png}
    \includegraphics[width=0.7\textwidth]{RatingDist_80.png}
	\caption{Distribution of rating pairs by common reviewers for item pairs.
Top figure shows distribution for item pairs having at least 40 common
reviewers. Bottom figure shows distribution for item pairs having at least 80
common reviewers.}
    \label{fig:RatingDist}
\end{figure}

The upper panel of Fig. \ref{fig:RatingDist} shows data for the $N \geq 40$
training set while the lower panel shows the same for the $N \geq 80$ training
set. In these density plots, for each rating pair, we randomly choose which is
Rating1 and which is Rating2. The two density plots present similar
distributions. These data, together with the similarity score histograms of Fig.
\ref{fig:Histograms}, suggest that the data for our two data sets are similar
and that large biases are not introduced by varying the number of common users,
$N$. At the same time, both rating distributions reveal significant correlations
between Rating1 and Rating2 as evidenced by the elongation in the $45^\circ$
direction. This is consistent with the data of Fig. \ref{fig:Histograms}, which
shows the mean $PearSim > 0$ for both sets. Further, these data indicate a
bias toward positive values among rating pairs compared to those of ratings in
general, which have an average value of zero by construction\footnote{See
discussion of bias subtraction in {\bf Data} section.}. This latter feature
indicates that items with large numbers of common users tend to be rated higher
by their common users than by their other users since we have already subtracted
the item's average rating thereby avoiding any simple popularity effects.

@@

\bibitem{Su2009}
    Xiaoyuan Su and Taghi M. Khoshgoftaar.
    ``A Survey of Collaborative Filtering Techniques.''
    \emph{Advances in Artificial Intelligence.} vol. 2009,
    Article ID 421425, 19 pages, 2009. doi:10.1155/2009/421425
\bibitem{Breese1998}
    Breese, John S., David Heckerman, and Carl Kadie, 
    ``Empirical analysis of predictive algorithms for collaborative filtering.'' 
    \emph{Proceedings of the Fourteenth conference on Uncertainty in artificial intelligence.}
    Morgan Kaufmann Publishers Inc., 1998.
\bibitem{Lemire1998}
    Daniel Lemire and Anna Maclachlan.
    ``Slope One Predictors for Online Rating-Based Collaborative Filtering.''
    \emph{SDM.} Vol. 5, 2005.
\bibitem{Koren2009}
    Koren, Yehuda, Robert Bell, and Chris Volinsky. 
    ``Matrix factorization techniques for recommender systems.''
    \emph{Computer} 42.8, 2009. 30-37.

>>>>

The parameter values $\mu_s$ and $\sigma_s$ are obtained by fitting the
distribution of $GoldSimRand$ values in the training set (see Fig.
\ref{fig:Histograms}). For example, in the $N \geq 80$ dataset we have $\mu_s =
0.27$ and $\sigma_s = 0.25$. The choice of $\mu_r$ and $\sigma_r$ is less clear.
Some insight can be gained, however, by examining the prior distribution of
ratings for our datasets, which are shown in Fig. \ref{fig:RatingDist}.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{RatingDist_40.png}
    \includegraphics[width=0.7\textwidth]{RatingDist_80.png}
	\caption{Distribution of rating pairs by common reviewers for item pairs.
Top figure shows distribution for item pairs having at least 40 common
reviewers. Bottom figure shows distribution for item pairs having at least 80
common reviewers. Rating1 and Rating2 are chosen in arbitrary order.}
    \label{fig:RatingDist}
\end{figure}

Fig. \ref{fig:RetingDist} shows density plots of rating pairs, for $N \geq 40$
and $N \geq 80$. These data, together with the similarity score histograms of Fig.
\ref{fig:Histograms}, suggest that the data for our two data sets are similar
and that large biases are not introduced by varying the number of common users,
$N$. At the same time, both rating distributions reveal significant correlations
between Rating1 and Rating2 as evidenced by the elongation in the $45^\circ$
direction. This is consistent with the data of Fig. \ref{fig:Histograms}, which
shows the mean $PearSim > 0$ for both sets. Further, these data indicate a
bias toward positive values among rating pairs compared to those of ratings in
general, which have an average value of zero by construction\footnote{See
discussion of bias subtraction in {\bf Data} section.}. This latter feature
indicates that items with large numbers of common users tend to be rated higher
by their common users than by their other users since we have already subtracted
the item's average rating thereby avoiding any simple popularity effects.

@@

\bibitem{Su2009}
    Xiaoyuan Su and Taghi M. Khoshgoftaar.
    ``A Survey of Collaborative Filtering Techniques.''
    \emph{Advances in Artificial Intelligence.} vol. 2009,
    Article ID 421425, 19 pages, 2009. doi:10.1155/2009/421425
\bibitem{Breese1998}
    John S. Breese, David Heckerman, and Carl Kadie, 
    ``Empirical analysis of predictive algorithms for collaborative filtering.'' 
    \emph{Proceedings of the Fourteenth conference on Uncertainty in artificial intelligence.}
    Morgan Kaufmann Publishers Inc., 1998.
\bibitem{Lemire1998}
    Daniel Lemire and Anna Maclachlan.
    ``Slope One Predictors for Online Rating-Based Collaborative Filtering.''
    \emph{SDM.} Vol. 5, 2005.
\bibitem{Koren2009}
    Yehuda Koren, Robert Bell, and Chris Volinsky. 
    ``Matrix factorization techniques for recommender systems.''
    \emph{Computer} 42.8, 2009. 30-37.
